### 2016-10-01 manuscript/manuscript.tex
Built corrected proof PDF

% Template for PLoS
% Version 3.1 February 2015
%
% To compile to pdf, run:
% latex plos.template
% bibtex plos.template
% latex plos.template
% latex plos.template
% dvipdf plos.template
%
% % % % % % % % % % % % % % % % % % % % % %
%
% -- IMPORTANT NOTE
%
% This template contains comments intended
% to minimize problems and delays during our production
% process. Please follow the template instructions
% whenever possible.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% Once your paper is accepted for publication,
% PLEASE REMOVE ALL TRACKED CHANGES in this file and leave only
% the final text of your manuscript.
%
% There are no restrictions on package use within the LaTeX files except that
% no packages listed in the template may be deleted.
%
% Please do not include colors or graphics in the text.
%
% Please do not create a heading level below \subsection. For 3rd level headings, use \paragraph{}.
%
% % % % % % % % % % % % % % % % % % % % % % %
%
% -- FIGURES AND TABLES
%
% Please include tables/figure captions directly after the paragraph where they are first cited in the text.
%
% DO NOT INCLUDE GRAPHICS IN YOUR MANUSCRIPT
% - Figures should be uploaded separately from your manuscript file.
% - Figures generated using LaTeX should be extracted and removed from the PDF before submission.
% - Figures containing multiple panels/subfigures must be combined into one image file before submission.
% For figure citations, please use "Fig." instead of "Figure".
% See http://www.plosone.org/static/figureGuidelines for PLOS figure guidelines.
%
% Tables should be cell-based and may not contain:
% - tabs/spacing/line breaks within cells to alter layout or alignment
% - vertically-merged cells (no tabular environments within tabular environments, do not use \multirow)
% - colors, shading, or graphic objects
% See http://www.plosone.org/static/figureGuidelines#tables for table guidelines.
%
% For tables that exceed the width of the text column, use the adjustwidth environment as illustrated in the example table in text below.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% -- EQUATIONS, MATH SYMBOLS, SUBSCRIPTS, AND SUPERSCRIPTS
%
% IMPORTANT
% Below are a few tips to help format your equations and other special characters according to our specifications. For more tips to help reduce the possibility of formatting errors during conversion, please see our LaTeX guidelines at http://www.plosone.org/static/latexGuidelines
%
% Please be sure to include all portions of an equation in the math environment.
%
% Do not include text that is not math in the math environment. For example, CO2 will be CO\textsubscript{2}.
%
% Please add line breaks to long display equations when possible in order to fit size of the column.
%
% For inline equations, please do not include punctuation (commas, etc) within the math environment unless this is part of the equation.
%
% % % % % % % % % % % % % % % % % % % % % % % %
%
% Please contact latex@plos.org with any questions.
%
% % % % % % % % % % % % % % % % % % % % % % % %

\documentclass[10pt,letterpaper]{article}
\usepackage[top=0.85in,left=2.75in,footskip=0.75in]{geometry}

% Use adjustwidth environment to exceed column width (see example table in text)
\usepackage{changepage}

% Use Unicode characters when possible
\usepackage[utf8]{inputenc}

% textcomp package and marvosym package for additional characters
\usepackage{textcomp,marvosym}

% fixltx2e package for \textsubscript
\usepackage{fixltx2e}

% amsmath and amssymb packages, useful for mathematical formulas and symbols
\usepackage{amsmath,amssymb}

% cite package, to clean up citations in the main text. Do not remove.
\usepackage{cite}

% Use nameref to cite supporting information files (see Supporting Information section for more info)
\usepackage{nameref}

% line numbers
\usepackage[right]{lineno}

% ligatures disabled
\usepackage{microtype}
\DisableLigatures[f]{encoding = *, family = * }

% rotating package for sideways tables
\usepackage{rotating}

% Remove comment for double spacing
%\usepackage{setspace}
%\doublespacing

%to handle url breaks in references
\usepackage{url}
\usepackage{hyperref}
\usepackage{breakurl}

% Text layout
\raggedright
\setlength{\parindent}{0.5cm}
\textwidth 5.25in
\textheight 8.75in

% Bold the 'Figure #' in the caption and separate it from the title/caption with a period
% Captions will be left justified
\usepackage[aboveskip=1pt,labelfont=bf,labelsep=period,justification=raggedright,singlelinecheck=off]{caption}

% Use the PLoS provided BiBTeX style
\bibliographystyle{resources/plos2015}

% Remove brackets from numbering in List of References
\makeatletter
\renewcommand{\@biblabel}[1]{\quad#1.}
\makeatother

% Leave date blank
\date{}

% Header and Footer with logo
\usepackage{lastpage,fancyhdr,graphicx}
\usepackage{epstopdf}
\pagestyle{myheadings}
\pagestyle{fancy}
\fancyhf{}
\lhead{\includegraphics[width=2.0in]{resources/PLOS-Submission.eps}}
\rfoot{\thepage/\pageref{LastPage}}
\renewcommand{\footrule}{\hrule height 2pt \vspace{2mm}}
\fancyheadoffset[L]{2.25in}
\fancyfootoffset[L]{2.25in}
\lfoot{\sf PLOS}

%% Include all macros below

\newcommand{\lorem}{{\bf LOREM}}
\newcommand{\ipsum}{{\bf IPSUM}}
% The below needed to build markdown documents containing lists with pandoc
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}
}

%% END MACROS SECTION

\begin{document}
\vspace*{0.35in}

% Title must be 250 characters or less.
% Please capitalize all terms in the title except conjunctions, prepositions, and articles.

\begin{flushleft}
{\Large
\textbf{Ten Simple Rules for Digital Data Storage}
}
\newline
% Insert author names, affiliations and corresponding author email (do not include titles, positions, or degrees).
\\
Edmund M. Hart $^{1 ,\ast}$,
Pauline Barmby $^{2 }$,
David LeBauer $^{3 }$,
François Michonneau $^{4,5 }$,
Sarah Mount $^{6 }$,
Patrick Mulrooney $^{11 }$,
Timothée Poisot $^{7 }$,
Kara H. Woo $^{8 }$,
Naupaka B. Zimmerman $^{9 }$,
Jeffrey W. Hollister $^{10 }$\\
\bigskip
\bf{1} University of Vermont, Department of Biology, Burlington
\\
\bf{2} University of Western Ontario, Department of Physics and Astronomy
\\
\bf{3} University of Illinois at Urbana-Champaign, National Center for Supercomputing Applications and Institute for
Genomic Biology
\\
\bf{4} University of Florida, iDigBio, Florida Museum of Natural History
\\
\bf{5} University of Florida, Whitney Laboratory for Marine Bioscience
\\
\bf{6} King's College London, Department of Informatics
\\
\bf{7} Université de Montréal, Département de Sciences Biologiques
\\
\bf{8} Washington State University, Center for Environmental Research, Education, and Outreach
\\
\bf{9} University of Arizona, School of Plant Sciences
\\
\bf{10} US Environmental Protection Agency, Atlantic Ecology Division
\\
\bf{11} University of California at San Diego, San Diego Supercomputer Center
\\
\bigskip
 $\ast$ E-mail: edmund.m.hart@gmail.com

\end{flushleft}

\linenumbers

\section*{Introduction}\label{introduction}
\addcontentsline{toc}{section}{Introduction}

Data is the central currency of science, but the nature of scientific
data has changed dramatically with the rapid pace of technology. This
change has led to the development of a wide variety of data formats,
dataset sizes, data complexity, data use cases, and data sharing
practices. Improvements in high-throughput DNA sequencing, sustained
institutional support for large sensor networks {[}1,2{]}, and sky
surveys with large-format digital cameras {[}3{]} have created massive
quantities of data. At the same time, the combination of increasingly
diverse research teams {[}4{]} and data aggregation in portals (e.g.,
for biodiversity data, \href{http://gbif.org/}{GBIF.org} or
\href{https://www.idigbio.org/portal/}{iDigBio}) necessitates increased
coordination among data collectors and institutions {[}5,6{]}. As a
consequence, ``data'' can now mean anything from petabytes of
information stored in professionally maintained databases, to
spreadsheets on a single computer, to handwritten tables in lab
notebooks on shelves. All remain important, but data curation practices
must continue to keep pace with the changes brought about by new forms
of data and new data collection and storage practices.

While much has been written about both the virtues of data sharing
{[}7,8{]} and the best practices to do so {[}9,10{]}, data storage has
received comparatively less attention. Proper storage is a prerequisite
to sharing, and indeed inadequate storage contributes to the phenomenon
of data decay or to ``data entropy,'' in which data, whether publicly
shared or not, becomes less accessible through time {[}11--14{]}. Best
practices for data storage often begin and end with this statement:
``Deposit your data in a community standard repository.'' This is good
advice, especially considering your data is most likely to be reused if
it is available on a community site. Community repositories can also
provide guidance for best practices. As an example, if you are archiving
sequencing data, a repository such as those run by the National Center
for Biotechnology Information (NCBI) (e.g., GenBank) not only provides a
location for data archival, but also encourages a set of practices
related to consistent data formatting and the inclusion of appropriate
metadata. However, data storage policies are highly variable between
repositories {[}15{]}. A data management plan utilizing best practices
across all stages of the data life cycle will facilitate transition from
local storage to repository {[}16{]}. Similarly, having such a plan can
facilitate transition from repository to repository if funding runs out
or requirements change. Good storage practices are important even (or
especially) in cases when data may not fit with an existing repository,
when only derived data products (versus raw data) are suitable for
archiving, or when an existing repository may have lax standards.

This article describes ten simple rules for digital data storage that
grew out of a long discussion among instructors for the Software and
Data Carpentry initiatives {[}17,18{]}. Software and Data Carpentry
instructors are scientists from diverse backgrounds who have encountered
a variety of data storage challenges and are active in teaching other
scientists best practices for scientific computing and data management.
Thus, this paper represents a distillation of collective experience, and
hopefully will be useful to scientists facing a variety of data storage
challenges. We additionally provide a glossary of common vocabulary for
readers who may not be familiar with particular terms.

\section*{Rule 1: Anticipate How Your Data Will Be
Used}\label{rule-1-anticipate-how-your-data-will-be-used}
\addcontentsline{toc}{section}{Rule 1: Anticipate How Your Data Will Be
Used}

One can avoid most of the troubles encountered during the analysis,
management, and release of data by having a clear roadmap of what to
expect before data acquisition starts. For instance:

\begin{itemize}
\tightlist
\item
  How will the raw data be received? Are they delivered by a machine or
  software, or typed in?
\item
  What is the format expected by the software used for analysis?
\item
  Is there a community standard format for this type of data?
\item
  How much data will be collected, and over what period of time?
\end{itemize}

The answers to these questions can range from simple cases (e.g.,
sequencing data stored in the FASTA format, which can be used ``as is''
throughout the analysis), to experimental designs involving multiple
instruments, each with its own output format and processing conventions.
Knowing the state in which the data needs to be at each step of the
analysis can help to (i) identify software tools to use in converting
between data formats, (ii) orient technological choices about how and
where the data should be stored, and (iii) rationalize the analysis
pipeline, making it more amenable to re-use {[}19{]}.

Also key is the ability to estimate the storage volume needed to store
the data, both during and after the analysis. The required strategy will
differ for datasets of varying size. Smaller datasets (e.g., a few
megabytes in size) can be managed locally with a simple data management
plan, whereas larger datasets (e.g., gigabytes to petabytes) will in
almost all cases require careful planning and preparation (Rule 10).

Lastly, early consideration and planning should be given to the metadata
of the project. A plan should be developed early as to what metadata
will be collected, and how it will be maintained and stored (Rule 7).
Also be sure to consider community software tools that can facilitate
metadata curation and repository submission. Examples in the biological
sciences include \emph{Morpho} for ecological metadata {[}20{]} and
\emph{mothur} {[}21{]} for submitting to NCBI's Sequence Read Archive.

\section*{Rule 2: Know Your Use Case}\label{rule-2-know-your-use-case}
\addcontentsline{toc}{section}{Rule 2: Know Your Use Case}

Well-identified use cases make data storage easier. Ideally, prior to
beginning data collection, researchers should be able to answer the
following questions:

\begin{itemize}
\tightlist
\item
  Should the raw data be archived (Rule 3)?
\item
  Should the data used for analysis be prepared once, or re-generated
  from the raw data each time (and what difference would this choice
  make for storage, computing requirements, and reproducibility)?
\item
  Can manual corrections be avoided in favor of programmatic or
  self-documenting approaches (e.g., Jupyter notebook or R markdown)?
\item
  How will changes to the data be tracked, and where will these tracked
  changes be logged?
\item
  Will the final data be released, and if so, in what format?
\item
  Are there restrictions or privacy concerns associated with the data
  (e.g., survey results with personally identifiable information
  {[}PII{]}, threatened species, or confidential business information)?
\item
  Will institutional validation be required prior to releasing the data?
\item
  Does the funding agency mandate data deposition in a publicly
  available archive, and if so, when, where, and under what license?
\item
  Does the target journal mandate data deposition?
\end{itemize}

None of these questions have universal answers, nor are they the only
questions to ask before starting data acquisition. But knowing the what,
when, and how of your use of the data will bring you close to a reliable
roadmap on how to handle data from acquisition through publication and
archival.

\section*{Rule 3: Keep Raw Data Raw}\label{rule-3-keep-raw-data-raw}
\addcontentsline{toc}{section}{Rule 3: Keep Raw Data Raw}

Since analytical and data processing procedures improve or otherwise
change over time, having access to the ``raw'' (unprocessed) data can
facilitate future re-analysis and analytical reproducibility. As
processing algorithms improve and computational power increases, new
analyses will be enabled that were not possible at the time of the
original work. If only derived data are stored, it can be difficult for
other researchers to confirm analytical results, to assess the validity
of statistical models, or to directly compare findings across studies.

Therefore, data should always be kept in raw format whenever possible
(within the constraints of technical limitations). In addition to being
the most appropriate way to ensure transparency in analysis, having the
data stored and archived in their original state gives a common point of
reference for derivative analyses. What constitutes sufficiently ``raw''
data is not always clear (e.g., ohms from a temperature sensor or images
of an Illumina sequencing flowcell are generally not archived after the
initial processing). Yet the spirit of this rule is that data should be
as ``pure'' as possible when they are stored. If derivations occur, they
should be documented by also archiving relevant code and intermediate
datasets.

A cryptographic hash (e.g., SHA or MD5) of the raw data should be
generated and distributed with the data. These hashes ensure that the
dataset has not suffered any silent corruption/manipulation while being
stored or transferred (see
\href{https://www.xsede.org/news/-/news/item/6390}{Internet2 Silent Data
Corruption}). For large enough datasets the likelihood of silent data
corruption is high. This technique has been widely used by many Linux
distributions to distribute images and has been very effective with
minimal effort.

\section*{Rule 4: Store Data in Open
Formats}\label{rule-4-store-data-in-open-formats}
\addcontentsline{toc}{section}{Rule 4: Store Data in Open Formats}

To maximize accessibility and long-term value, it is preferable to store
data in formats that have freely available specifications. The
appropriate file type will depend on the data being stored (e.g.,
numeric measurements, text, images, video), but the key idea is that
accessing data should not require proprietary software, hardware, or
purchase of a commercial license. Proprietary formats change,
maintaining organizations go out of business, and changes in license
fees make access to data in proprietary formats unaffordable and risky
for end-users. Examples of open data formats include comma-separated
values (CSV) for tabular data, hierarchical data format (HDF) {[}22{]}
and NetCDF {[}23{]} for hierarchically structured scientific data,
portable network graphics (PNG) for images, KML (or other Open
Geospatial Consortium {[}OGC{]} format) for spatial data, and extensible
markup language (XML) for documents. Examples of closed formats include
DWG for AutoCAD drawings, Photoshop document (PSD) for bitmap images,
Windows Media Audio (WMA) for audio recording files, and Microsoft Excel
(XLS) for tabular data. Even if day-to-day processing uses closed
formats (e.g., due to software requirements), data being stored for
archival purposes should be stored in open formats. This is generally
not prohibitive; most closed-source software products enable users to
export data to an open format.

Not only should data be stored in an open format, but it should also be
stored in a format that computers can easily use for processing. This is
especially crucial as datasets become larger. Making data easily usable
is best achieved by using standard data formats that have open
specifications (e.g., CSV, XML, JSON, HDF5), or by using databases. Such
data formats can be handled by a variety of programming languages, as
efficient and well-tested libraries for parsing them are typically
available. These standard data formats also ensure interoperability,
facilitate re-use, and reduce the chances of data loss or mistakes being
introduced during conversion between formats. Examples of
machine-readable open formats that would not be easy to process include
data included in the text of a PDF file, or scanned images of tabular
data from a paper source.

\section*{Rule 5: Data Should Be Structured for
Analysis}\label{rule-5-data-should-be-structured-for-analysis}
\addcontentsline{toc}{section}{Rule 5: Data Should Be Structured for
Analysis}

To take full advantage of data, it can be useful for it to be structured
in a way that makes use, interpretation, and analysis easy. One such
structure for data stores each variable as a column, each observation as
a row, and each type of observational unit as a table (Fig. 1). The
technical term for this structure is ``Codd's 3rd normal form'', but it
has been made more accessible as the concept of tidy data {[}24{]}. When
data is organized in this way, the duplication of information is reduced
and it is easier to subset or summarize the dataset to include the
variables or observations of interest.

\textbf{Figure 1}: Example of an untidy dataset (A) and its tidy
equivalent (B). Dataset A is untidy because it mixes observational units
(species, location of observations, measurements about individuals), the
units are mixed and listed with the observations, more than one variable
is listed (both latitude and longitude for the coordinates, and genus
and species for the species names), and several formats are used in the
same column for dates and geographic coordinates. Dataset B is an
example of a tidy version of dataset A that reduces the amount of
information that is duplicated in each row, limiting chances of
introducing mistakes in the data. By having species in a separate table,
they can be identified uniquely using the Taxonomic Serial Number (TSN)
from the Integrated Taxonomic Information System (ITIS), and it makes it
easy to add information about the classification of these species. It
also allows researchers to edit the taxonomic information independently
from the table that holds the measurements about the individuals. Unique
values for each observational unit facilitate the programmatic
combination of information using ``join'' operations. With this example,
if the focus of the study for which these data were collected is based
upon the size measurements of the individuals (weight and length),
information about ``where'', ``when'', and ``what'' animals were
measured can be considered meta-data. Using the tidy format makes this
distinction clearer.

One axiom about the structure of data and code holds that one should
``write code for humans, write data for computers'' {[}25{]}. When data
can be easily imported and manipulated using familiar software (whether
via a scripting language, a spreadsheet, or any other computer program
that can import these common files), data becomes easier to re-use.
Furthermore, having the source code for the software doing the analysis
available provides provenance for how the data is processed and
analyzed. This makes analysis more transparent, since all assumptions
about the structure of the data are implicitly stated in the source
code. This also enables extraction of the analyses performed, their
reproduction, and their modification.

Interoperability is facilitated when variable names are mapped to
existing data standards. For instance, for biodiversity data, the
\href{http://www.tdwg.org/standards/450/}{Darwin Core Standard} provides
a set of terms that describe observations, specimens, samples, and
related information for a taxa. For earth science and ecosystem models
and data, the \href{https://cfconventions.org}{Climate Forecasting
Conventions} are widely adopted, such that a large ecosystem of software
and data products exist to reduce the technical burden of reformatting
and reusing large and complex data. Because each term in such standards
is clearly defined and documented, each dataset can use the terms
consistently; this facilitates data sharing across institutions,
applications, and disciplines. With machine-readable,
standards-compliant data, it becomes easier to build an Application
Programming Interface (API) to query the dataset and retrieve a subset
of interest, as outlined in Rule 10.

\section*{Rule 6: Data Should Be Uniquely
Identifiable}\label{rule-6-data-should-be-uniquely-identifiable}
\addcontentsline{toc}{section}{Rule 6: Data Should Be Uniquely
Identifiable}

To aid reproducibility, the data used in a scientific publication should
be uniquely identifiable. Ideally, datasets should have a unique
identifier such as a Digital Object Identifier (DOI), Archival Resource
Key (ARK), or a persistent URL (PURL). An increasing number of online
services, such as \href{http://figshare.com/}{Figshare},
\href{http://zenodo.org}{Zenodo}, or
\href{http://www.dataone.org}{DataOne}, are able to provide these.
Institutional initiatives also exist and are known to your local
librarians. Some repositories may require specific identifiers, and
these could change with time. For instance, NCBI sequence data will in
the future only be identified by ``accession.version'' IDs. The ``GI''
identifiers (in use since 1994) will be retired in late 2016 {[}26{]}.

Even as identifier standards may change over time, datasets can evolve
over time as well. In order to distinguish between different versions of
the same data, each dataset should have a distinct name, which includes
a version identifier. A simple way to do this is to use date stamps as
part of the dataset name. Using the ISO 8601 standard avoids regional
ambiguities: it mandates the date format \texttt{YYYY-MM-DD} (i.e.~from
largest time unit to smallest). For example, the date ``February 1,
2015,'' while written as \texttt{01-02-2015} in the UK and
\texttt{02-01-2015} in the US, is unambiguous (\texttt{2015-02-01})
under this standard.

Semantic versioning is a richer approach to solving the same problem
{[}27{]}. The CellPack datasets are an example of this {[}28{]}. A
semantic version number takes the form: \texttt{Major.Minor.Patch} e.g.,
\texttt{0.2.7}. The \emph{major version} numbers should be incremented
(or bumped) when a dataset scheme has been updated, or some other change
is made that is not compatible with previous versions of the data with
the same major version number. This means that an experiment using
version \texttt{1.0.0} of the dataset may not run on version
\texttt{2.0.0} without changes to the data analysis. The \emph{minor
version} should be bumped when a change has been made that is compatible
with older versions of the data with the same major version. This means
that any analysis that can be performed on version \texttt{1.0.0} of the
data is repeatable with version \texttt{1.1.0} of the data. For example,
adding a new year in a temporal survey will result in a bump in the
minor version. The \emph{patch version} number is bumped when typos or
bugs have been fixed. For example version \texttt{1.0.1} of a dataset
may fix a typo in version \texttt{1.0.0}.

\section*{Rule 7: Link Relevant
Metadata}\label{rule-7-link-relevant-metadata}
\addcontentsline{toc}{section}{Rule 7: Link Relevant Metadata}

Metadata is the contextual information required to interpret data (Fig.
1) and should be clearly defined and tightly integrated with data. The
importance of metadata for context, reusability, and discovery has been
written about at length in guides for data management best practices
{[}9,13,29{]}.

Metadata should be as comprehensive as possible, using standards and
conventions of a discipline, and should be machine-readable. Metadata
should always accompany a dataset, wherever it is stored, but the best
way to do this depends on the format of the data. Text files can contain
metadata in well-defined text files such as XML or JSON. Some file
formats are self-documenting; for example, NetCDF, HDF5, and many image
file formats allow for embedded metadata {[}22,23{]}. In a relational
database, metadata tables can be clearly labeled and linked to the data.
Ideally, a schema will be provided that also shows the linkages between
data tables and metadata tables. Another---simpler---scenario is a set
of flat (non-hierarchical) text files---in this case a semantically
versioned, compressed archive should be created that includes metadata.

Whatever format is used for archiving, the goal should be to make the
link between metadata and data as clear as possible. The best approach
is dependent on the archiving plan used, but even if the dataset is
archived solely for personal use, metadata will provide crucial context
for future reuse.

\section*{Rule 8: Adopt the Proper Privacy
Protocols}\label{rule-8-adopt-the-proper-privacy-protocols}
\addcontentsline{toc}{section}{Rule 8: Adopt the Proper Privacy
Protocols}

In datasets for which privacy is important, be sure to have a plan in
place to protect data confidentiality. You should consider the different
data stakeholders when developing privacy protocols for your data
storage. These stakeholders include funding agencies, human subjects or
entities, collaborators, and yourself. Both the National Science
Foundation (NSF) and National Institutes of Health (NIH) have data
sharing policies in their grant guidelines to prevent sharing personally
identifiable information, and to anonymize data on human subjects.

In small datasets, a lookup table (protecting PII by removing it and
replacing it with a unique ID that maps to the sensitive data in an
external dataset) is enough to anonymize a minimal amount of personal
information. Hashing techniques are susceptible to a number of attacks,
and all hashed data can eventually be determined. Famously, New York
City officials shared what they thought was anonymized data on cab
drivers and over 173 million cab rides. However, it was quickly
recognized that the city anonymized the data with a simple MD5 hashing
scheme and all 20 GB of data were de-anonymized in a matter of hours
{[}30{]}. This type of error can be prevented by asking a trusted
colleague or security personnel to try to ``crack'' anonymised data
before releasing it publicly. Often the person who has produced the data
is not well placed to check the fine details of their own security
procedures. If possible, the best solution is to remove any sensitive
data that is not required from the dataset prior to distribution.

In more problematic cases, the data itself allows identifiability: this
is the case with human genomic data that map directly onto a subject's
identity {[}31{]}. Methods for dealing with these complex issues at the
intersection of data storage and privacy are rapidly evolving, and
include storing changes against a reference genome to help with privacy
and reduce overall data volumes {[}32,33{]}, and/or bringing computation
to data storage facilities instead of vice versa {[}34{]}. Having a plan
for privacy before data is acquired is important, because it can
determine or limit how data will be stored.

\section*{Rule 9: Have a Systematic Backup
Scheme}\label{rule-9-have-a-systematic-backup-scheme}
\addcontentsline{toc}{section}{Rule 9: Have a Systematic Backup Scheme}

Every storage medium can fail, and every failure can result in loss of
data. Researchers should therefore back up data at all stages of the
research process. Data stored on local computers or institutional
servers during the collection and analysis phases of a project should be
backed up to other locations to protect against data loss. No backup
system is failsafe (see the stories of the
\href{https://www.insidehighered.com/news/2014/05/16/dedoose-crash-shows-dangers-handing-data-cloud-services}{Dedoose
crash} and the
\href{http://thenextweb.com/media/2012/05/21/how-pixars-toy-story-2-was-deleted-twice-once-by-technology-and-again-for-its-own-good/}{near
deletion of Toy Story 2}), so more than one backup system should be
used. Ideally you should have two on-site copies (such as on a computer,
an external hard drive, or a tape) and one off-site copy (e.g., cloud
storage){[}35{]}, with care taken to ensure that the off-site copy is as
secure as the on-site copies. Keeping backups in multiple locations
additionally protects against data loss due to theft or natural
disasters.

Researchers should test their backups regularly to ensure that they are
functioning properly. Common reasons for backup failure include:

\begin{itemize}
\tightlist
\item
  faulty backup software
\item
  incorrect configuration (e.g., not backing up sub-directories)
\item
  encryption (e.g., someone encrypted the backups but later lost the
  password to decrypt them)
\item
  media errors
\end{itemize}

Consider the backup plan of your selected data repository before
publishing your data and if possible, find out about the long-term
storage plans of the repository. Many repositories mirror the data they
host on multiple machines. Are there plans in place to keep data
available if the organization that manages the repository dissolves?

\section*{Rule 10: The Location and Method of Data Storage Depend on How
Much Data You
Have}\label{rule-10-the-location-and-method-of-data-storage-depend-on-how-much-data-you-have}
\addcontentsline{toc}{section}{Rule 10: The Location and Method of Data
Storage Depend on How Much Data You Have}

The storage method you should choose depends on the size and nature of
your data, the cost of storage and access over time, the time it takes
to transfer the data, how the data will be used, and any privacy
concerns. Data is increasingly generated in the range of many terabytes
by environmental sensors, satellites, automated analytical tools,
simulation models, and nucleic acid sequencers. Even larger
data-generating machines, like the Large Hadron Collider (LHC) and the
Large Scale Synoptic Survey Telescope (LSST), generate many TB per day,
rapidly accumulating to petabyte (PB) scale over the course of any
particular study. While the cost of storage continues to decrease, the
volume of data to be stored impacts the choice of storage methods and
locations: for large datasets it is necessary to balance the cost of
storage with the time of access and costs of re-generating the data.
With new commercial cloud offerings (e.g., Amazon S3) the cost of
retrieving the data might exceed the cost of analysis or of
re-generating the data from scratch.

When data takes too long to transfer or is costly to store, it can
become more efficient to use a computer system for analysis that can
directly access and use the data in place instead of first transferring
it to a local machine. Inactive data can be put in longer-term storage;
this is less expensive, but can take longer to retrieve. Some storage
systems automatically migrate ``stale'' files to longer-term storage.
Alternatively, some computing can be done ``in the database'' or ``on
disk'' via database query languages (e.g., SQL, MapReduce) that perform
basic arithmetic, or via the use of procedural languages (e.g., R,
Python, C) embedded in the database server. Modern database technologies
such as HDFS and Spark allow these computations to be done on data of
almost any size. When data is larger than locally available RAM, it can
be handled by conducting analyses on a ``big memory'' node, which most
high-performance computing centers have deployed. Relying on tight
software/hardware integration, these can allow for the analysis of
datasets around 1--4 TB in size. This allows the user to read in and use
a large dataset without special tools.

If you regularly only need access to a small subset of your data or need
to share it with many collaborators, an online API (Application
Programming Interface) might be a good solution. Using this method, many
users can send requests to a online service that can subset the data,
perform in-database computation, and return smaller volumes of data as
specific slices. Tools based on online services make it easier to find
and download data, and they facilitate analysis via reproducible
scripts. However, they can also lead to excessive and careless abuse of
resources without proper safeguards in place. The time required to
re-download and re-compute results can be reduced by ``caching''.
Caching stores copies of downloads and generated files that are
recognized when the same script is run multiple times.

\section*{Further Reading and
Resources}\label{further-reading-and-resources}
\addcontentsline{toc}{section}{Further Reading and Resources}

Digital data storage is a vast topic; the references given here and
elsewhere in this paper provide some starting points for interested
readers. For beginning users of scientific data,
\href{http://datacarpentry.org}{Data Carpentry} offers workshops and
resources on data management and analysis, as do the DataONE education
modules {[}36{]}. For librarians and others who are responsible for data
archiving, Data Curation Profiles {[}37{]} may be of interest.

\section*{Glossary}\label{glossary}
\addcontentsline{toc}{section}{Glossary}

\subsection*{Projects and Initiatives}\label{projects-and-initiatives}
\addcontentsline{toc}{subsection}{Projects and Initiatives}

\begin{itemize}
\item
  \textbf{Global Biodiversity Information Facility} (GBIF,
  \url{http://www.gbif.org}) provides an international open data
  infrastructure to publish and disseminate biodiversity information.
\item
  \textbf{Integrated Digitized Biocollections} (iDigBio,
  \url{https://www.idigbio.org}) is a project funded by the National
  Science Foundation that facilitates the digitization of natural
  history collections, and provides data and images for biological
  specimens.
\item
  \textbf{Integrated Taxonomic Information System} (ITIS,
  \url{http://www.itis.gov}) is an international partnership of
  governmental organizations that aims at providing authoritative
  taxonomic information for plants, animals, fungi, and microbes.
\end{itemize}

\subsection*{File formats}\label{file-formats}
\addcontentsline{toc}{subsection}{File formats}

\begin{itemize}
\item
  \textbf{Comma-Separated Values} (CSV) and \textbf{Tab-Separated
  Values} (TSV) are plain text file formats used to store tabular data,
  in which each row is represented by a line in the file, and each field
  (column) is separated by a comma (for CSV) or by the tab character
  (for TSV).
\item
  \textbf{FASTA} is a simple and widely used file format used to
  represent sequences of nucleotides or amino acids in plain text,
  making it easy to manipulate these programmatically.
\item
  \textbf{Hierarchical Data Format} (HDF5) is an open-source binary file
  format designed to store large amounts of data (and their associated
  metadata) by providing a hierarchical structure that could be compared
  to how a hard drive is organized with directories and files. It is
  maintained by the non-profit HDF Group, a spin-off of the National
  Center for Supercomputing Applications (NCSA).
\item
  \textbf{JavaScript Object Notation} (JSON) is a plain text file format
  typically used to store arbitrarily structured data in the form of
  keys and values. It can be used to store non-relational databases, as
  it does not rely on a tabular data format. In many respects, it has
  been replacing XML.
\item
  \textbf{Network Common Data Form} (NetCDF) is an open-source binary
  file format designed to store large datasets in array-oriented
  scientific data, typically used in the geosciences. It is maintained
  by Unidata, a non-profit member of the University Corporation for
  Atmospheric Research (UCAR), which is funded by the National Science
  Foundation.
\item
  \textbf{Extensible Markup Language} (XML) is a markup language and the
  file format used to store documents written with it. It is used to
  represent arbitrary data structures and is both human and
  machine-readable.
\end{itemize}

\subsection*{Programming and
algorithms}\label{programming-and-algorithms}
\addcontentsline{toc}{subsection}{Programming and algorithms}

\begin{itemize}
\item
  \textbf{Web Application Programming Interface} (API) provide ways to
  programmatically query databases through the internet. They notably
  allow users to retrieve and work with a small slice of a large
  dataset.
\item
  \textbf{Hadoop Distributed File System} (HDFS) is a Java-based file
  system in which data is stored in small chunks across multiple
  redundant nodes.
\item
  \textbf{MapReduce} is a style of programming designed to work with
  large datasets in parallel computing environments. Such programs are
  composed of a \textbf{map} procedure in which the dataset is sliced
  into several pieces, and a \textbf{reduce} procedure in which summary
  operations are then applied to each of the slices.
\item
  \textbf{Secure Hash Algorithm 2} (SHA-2) is a family of Secure Hashing
  Algorithms used in cryptographic analysis, often to verify the
  integrity of a file. A cryptographic hash function converts a
  ``message'' (e.g., passwords, file content) into an encrypted value.
  Cryptographic hash functions are easy to compute from the message, but
  it should be impossible to recover the message from the output, and
  any modifications to the message should also modify the output. The
  SHA algorithms are often used in preference to similar tools such as
  MD5 (mentioned in Rule 3 and 8), which are no longer secure. All
  hashing algorithms are vulnerable to brute force attacks. Key
  Derivation Function (KDF) implementations like BCrypt and PBKDF2 are
  considered significantly more secure, but by design more costly to
  compute.
\item
  Apache \textbf{Spark} is an open-source computing platform for
  querying large datasets in memory, in contrast to on--disk--based
  methods like MapReduce.
\item
  \textbf{Structured Query Language} (SQL) is a programming language
  used to interact with relational database management systems.
\end{itemize}

\subsection*{Hardware}\label{hardware}
\addcontentsline{toc}{subsection}{Hardware}

\begin{itemize}
\tightlist
\item
  \textbf{mega-, giga-, tera-, peta-bytes} are units of digital
  information and are used to measure the size of datasets or the
  storage media. Originally a byte was the minimum amount of memory
  needed to store a single character of text in a computer. The prefixes
  mega-, giga-, tera-, and peta- refer to the international system of
  units for the multiple of the unit, and correspond to \(10^{6}\),
  \(10^{9}\), \(10^{12}\), and \(10^{15}\), abbreviated M, G, T, and P,
  respectively.
\end{itemize}

\subsection*{Persistent identifiers}\label{persistent-identifiers}
\addcontentsline{toc}{subsection}{Persistent identifiers}

\begin{itemize}
\item
  \textbf{Archival Resource Key} (ARK) identifiers are URLs designed to
  support long-term access to information online.
\item
  \textbf{Digital Object Identifier} (DOI) provides unique and
  persistent identifiers for electronic documents (in particular,
  journal articles and datasets) on the internet. The uniqueness of the
  identifiers is guaranteed by a central registry. By dissociating the
  identifier and the location of the document (i.e., the URL), the DOI
  can remain fixed even if the location of the digital object it is
  pointing to changes.
\item
  \textbf{Persistent Uniform Resource Locator} (PURL) is a URL used to
  redirect to the location of an electronic object on the internet. DOI
  and ARK are examples of implementations of PURL.
\item
  \textbf{Uniform Resource Locator} (URL) gives the location of an
  object on the World Wide Web; the most familiar type of URL is a
  website address.
\end{itemize}

\section*{Acknowledgements}\label{acknowledgements}
\addcontentsline{toc}{section}{Acknowledgements}

We would like to thank G. Wilson and the Software Carpentry instructor
community for providing the initial inspiration for this paper and the
discussion held at \url{https://github.com/swcarpentry/site/issues/797}.
We thank K. Hinsen, G. Pugh, and W. Turner for comments on the draft
manuscript. Contributions from J. W. Hollister, an employee of the US
Environmental Protection Agency (EPA) have not been subjected to the
Agency's peer and administrative review. Therefore, the conclusions and
opinions drawn are solely those of the author(s) and are not necessarily
the views of the Agency. Mention of trade names or commercial products
does not constitute endorsement or recommendation for use. S. Mount
would like to thank Dr.~Laurence Tratt for reviewing an early draft of
this paper.

\newpage

\section*{References}\label{references}
\addcontentsline{toc}{section}{References}

\bibdata{resources/manuscript.bib}

\hyperdef{}{ref-Reid2014}{\label{ref-Reid2014}}
1. Reid JG, Carroll A, Veeraraghavan N, Dahdouli M, Sundquist A, English
A, et al. Launching genomics into the cloud: deployment of Mercury, a
next generation sequence analysis pipeline. BMC bioinformatics. 2014;15:
30.
\href{http://doi.org/10.1186/1471-2105-15-30}{doi:10.1186/1471-2105-15-30}

\hyperdef{}{ref-Hampton2013}{\label{ref-Hampton2013}}
2. Hampton SE, Strasser C, Tewksbury JJ, Gram WK, Budden AE, Batcheller
AL, et al. Big data and the future of ecology. Frontiers in Ecology and
the Environment. 2013; 130312142848005.
\href{http://doi.org/10.1890/120103}{doi:10.1890/120103}

\hyperdef{}{ref-Eisenstein2011}{\label{ref-Eisenstein2011}}
3. Eisenstein DJ, others. SDSS-III: Massive Spectroscopic Surveys of the
Distant Universe, the Milky Way, and Extra-Solar Planetary Systems. The
Astronomical Journal. 2011;142: 72.

\hyperdef{}{ref-Adams2012}{\label{ref-Adams2012}}
4. Adams J. Collaborations: The rise of research networks. Nature.
Nature Publishing Group, a division of Macmillan Publishers Limited. All
Rights Reserved. 2012;490: 335--6.
\href{http://doi.org/10.1038/490335a}{doi:10.1038/490335a}

\hyperdef{}{ref-Fraser2013}{\label{ref-Fraser2013}}
5. Fraser LH, Henry HA, Carlyle CN, White SR, Beierkuhnlein C, Cahill
JF, et al. Coordinated distributed experiments: an emerging tool for
testing global hypotheses in ecology and environmental science.
Frontiers in Ecology and the Environment. Ecological Society of America;
2013;11: 147--155.
\href{http://doi.org/10.1890/110279}{doi:10.1890/110279}

\hyperdef{}{ref-Robertson2014}{\label{ref-Robertson2014}}
6. Robertson MAG Tim AND Döring. The GBIF integrated publishing toolkit:
Facilitating the efficient publishing of biodiversity data on the
internet. PLoS ONE. Public Library of Science; 2014;9: e102623.
\href{http://doi.org/10.1371/journal.pone.0102623}{doi:10.1371/journal.pone.0102623}

\hyperdef{}{ref-Wolkovich2012}{\label{ref-Wolkovich2012}}
7. Wolkovich EM, Regetz J, O'Connor MI. Advances in global change
research require open science by individual researchers. Global Change
Biology. 2012;18: 2102--2110.
\href{http://doi.org/10.1111/j.1365-2486.2012.02693.x}{doi:10.1111/j.1365-2486.2012.02693.x}

\hyperdef{}{ref-Roche2014}{\label{ref-Roche2014}}
8. Roche DG, Lanfear R, Binning SA, Haff TM, Schwanz LE, Cain KE, et al.
Troubleshooting public data archiving: suggestions to increase
participation. PLoS biology. 2014;12: e1001779.
\href{http://doi.org/10.1371/journal.pbio.1001779}{doi:10.1371/journal.pbio.1001779}

\hyperdef{}{ref-White2013}{\label{ref-White2013}}
9. White E, Baldridge E, Brym Z, Locey K, McGlinn D, Supp S. Nine simple
ways to make it easier to (re)use your data. Ideas in Ecology and
Evolution. 2013;6: 1--10.
\href{http://doi.org/10.4033/iee.2013.6b.6.f}{doi:10.4033/iee.2013.6b.6.f}

\hyperdef{}{ref-Goodman2014}{\label{ref-Goodman2014}}
10. Goodman A, Pepe A, Blocker AW, Borgman CL, Cranmer K, Crosas M, et
al. Ten simple rules for the care and feeding of scientific data. PLoS
computational biology. 2014;10: e1003542.
\href{http://doi.org/10.1371/journal.pcbi.1003542}{doi:10.1371/journal.pcbi.1003542}

\hyperdef{}{ref-Pepe2014}{\label{ref-Pepe2014}}
11. Pepe A, Goodman A, Muench A, Crosas M, Erdmann C. How Do Astronomers
Share Data? Reliability and Persistence of Datasets Linked in AAS
Publications and a Qualitative Study of Data Practices among US
Astronomers. Golden AA-J, editor. PLoS ONE. 2014;9: e104798.
\href{http://doi.org/10.1371/journal.pone.0104798}{doi:10.1371/journal.pone.0104798}

\hyperdef{}{ref-Vines2014}{\label{ref-Vines2014}}
12. Vines TH, Albert AYK, Andrew RL, Débarre F, Bock DG, Franklin MT, et
al. The availability of research data declines rapidly with article age.
Current biology : CB. Elsevier; 2014;24: 94--7.
\href{http://doi.org/10.1016/j.cub.2013.11.014}{doi:10.1016/j.cub.2013.11.014}

\hyperdef{}{ref-Michener2012}{\label{ref-Michener2012}}
13. Michener WK, Jones MB. Ecoinformatics: supporting ecology as a
data-intensive science. Trends in ecology \& evolution. 2012;27: 85--93.
\href{http://doi.org/10.1016/j.tree.2011.11.016}{doi:10.1016/j.tree.2011.11.016}

\hyperdef{}{ref-Michener1997}{\label{ref-Michener1997}}
14. Michener WK, Brunt JW, Helly JJ, Kirchner TB, Stafford SG.
Nongeospatial metadata for the ecological sciences. Ecological
Applications. Eco Soc America; 1997;7: 330--342.

\hyperdef{}{ref-Marcial2010}{\label{ref-Marcial2010}}
15. Marcial LH, Hemminger BM. Scientific data repositories on the Web:
An initial survey. Journal of the American Society for Information
Science and Technology. 2010;61: 2029--2048.
\href{http://doi.org/10.1002/asi.21339}{doi:10.1002/asi.21339}

\hyperdef{}{ref-Michener2015}{\label{ref-Michener2015}}
16. Michener WK. Ten simple rules for creating a good data management
plan. PLoS computational biology. Public Library of Science; 2015;11:
e1004525.

\hyperdef{}{ref-Wilson2014}{\label{ref-Wilson2014}}
17. Wilson G. Software Carpentry: lessons learned. F1000Research.
2014;3: 62.
\href{http://doi.org/10.12688/f1000research.3-62.v1}{doi:10.12688/f1000research.3-62.v1}

\hyperdef{}{ref-Teal2015}{\label{ref-Teal2015}}
18. Teal TK, Cranston Ka, Lapp H, White E, Wilson G, Ram K, et al. Data
Carpentry: Workshops to Increase Data Literacy for Researchers.
International Journal of Digital Curation. 2015;10: 135--143.
\href{http://doi.org/10.2218/ijdc.v10i1.351}{doi:10.2218/ijdc.v10i1.351}

\hyperdef{}{ref-Roche2015}{\label{ref-Roche2015}}
19. Roche DG, Kruuk LEB, Lanfear R, Binning SA. Public Data Archiving in
Ecology and Evolution: How Well Are We Doing? PLoS Biology. 2015;13:
e1002295.
\href{http://doi.org/10.1371/journal.pbio.1002295}{doi:10.1371/journal.pbio.1002295}

\hyperdef{}{ref-higgins2002}{\label{ref-higgins2002}}
20. Higgins D, Berkley C, Jones M. Managing heterogeneous ecological
data using morpho {[}Internet{]}. IEEE Comput. Soc; 2002. pp. 69--76.
\href{http://doi.org/10.1109/SSDM.2002.1029707}{doi:10.1109/SSDM.2002.1029707}

\hyperdef{}{ref-schloss2009}{\label{ref-schloss2009}}
21. Schloss PD, Westcott SL, Ryabin T, Hall JR, Hartmann M, Hollister
EB, et al. Introducing mothur: Open-source, platform-independent,
community-supported software for describing and comparing microbial
communities. Applied and environmental microbiology. Am Soc Microbiol;
2009;75: 7537--7541.

\hyperdef{}{ref-koziol1998hdf5}{\label{ref-koziol1998hdf5}}
22. Koziol Q, Matzke R. Hdf5: A new generation of hdf: Reference manual
and user guide. National Center for Supercomputing Applications,
Champaign, Illinois, USA, http://hdf ncsa uiuc edu/nra/HDF5. 1998;

\hyperdef{}{ref-rew1990netcdf}{\label{ref-rew1990netcdf}}
23. Rew R, Davis G. NetCDF: An interface for scientific data access.
Computer Graphics and Applications, IEEE. IEEE; 1990;10: 76--82.

\hyperdef{}{ref-Wickham2014tidy}{\label{ref-Wickham2014tidy}}
24. Wickham H. Tidy Data. Journal of Statistical Software. 2014;59:
1--23. Available: \url{http://www.jstatsoft.org/v59/i10}

\hyperdef{}{ref-buffalo2015bioinformatics}{\label{ref-buffalo2015bioinformatics}}
25. Buffalo V. Bioinformatics data skills: Reproducible and robust
research with open source tools. O'Reilly Media, Inc. 2015.

\hyperdef{}{ref-ncbi}{\label{ref-ncbi}}
26. NCBI. NCBI is phasing out sequence gIs - use accession.Version
instead!
\url{http://www.ncbi.nlm.nih.gov/news/03-02-2016-phase-out-of-GI-numbers/};

\hyperdef{}{ref-semver2014}{\label{ref-semver2014}}
27. Preston-Werner T. Semantic Versioning 2.0.0.
\url{http://semver.org}; 2014.

\hyperdef{}{ref-johnson2014cellpack}{\label{ref-johnson2014cellpack}}
28. Johnson GT, Goodsell DS, Autin L, Forli S, Sanner MF, Olson AJ. 3D
molecular models of whole HIV-1 virions generated with cellPACK. Faraday
Discuss. The Royal Society of Chemistry; 2014;169: 23--44.
\href{http://doi.org/10.1039/C4FD00017J}{doi:10.1039/C4FD00017J}

\hyperdef{}{ref-Strasser2012}{\label{ref-Strasser2012}}
29. Strasser C, Cook R, Michener W, Budden A. Primer on Data Management
: What you always wanted to know {[}Internet{]}. California Digital
Libraries; 2012.
\href{http://doi.org/10.5060/D2251G48}{doi:10.5060/D2251G48}

\hyperdef{}{ref-Goodin}{\label{ref-Goodin}}
30. Goodin D. Poorly anonymized logs reveal NYC cab drivers' detailed
whereabouts {[}Internet{]}. 2015. Available:
\url{http://arstechnica.com/tech-policy/2014/06/poorly-anonymized-logs-reveal-nyc-cab-drivers-detailed-whereabouts/}

\hyperdef{}{ref-Homer2008}{\label{ref-Homer2008}}
31. Homer N, Szelinger S, Redman M, Duggan D, Tembe W, Muehling J, et
al. Resolving individuals contributing trace amounts of DNA to highly
complex mixtures using high-density SNP genotyping microarrays. PLoS
genetics. Public Library of Science; 2008;4: e1000167.
\href{http://doi.org/10.1371/journal.pgen.1000167}{doi:10.1371/journal.pgen.1000167}

\hyperdef{}{ref-Kahn2011}{\label{ref-Kahn2011}}
32. Kahn SD. On the future of genomic data. Science (New York, NY).
2011;331: 728--9.
\href{http://doi.org/10.1126/science.1197891}{doi:10.1126/science.1197891}

\hyperdef{}{ref-wandelt2014trends}{\label{ref-wandelt2014trends}}
33. Wandelt S, Bux M, Leser U. Trends in genome compression. Current
Bioinformatics. Bentham Science Publishers; 2014;9: 315--326.

\hyperdef{}{ref-Gaye2014}{\label{ref-Gaye2014}}
34. Gaye A, Marcon Y, Isaeva J, LaFlamme P, Turner A, Jones EM, et al.
DataSHIELD: taking the analysis to the data, not the data to the
analysis. International journal of epidemiology. 2014;43: 1929--44.
\href{http://doi.org/10.1093/ije/dyu188}{doi:10.1093/ije/dyu188}

\hyperdef{}{ref-briney2013}{\label{ref-briney2013}}
35. Briney K. Rule of 3. DATA AB INITIO.
\url{http://dataabinitio.com/?p=320}; 2013.

\hyperdef{}{ref-Dataone2012}{\label{ref-Dataone2012}}
36. Henkel H, Hutchison V, Strasser C, Rebich Hespanha S, Vanderbilt K,
Wayne L, et al. DataONE Education Modules.
\url{https://www.dataone.org/education-modules}; 2012.

\hyperdef{}{ref-Witt2009}{\label{ref-Witt2009}}
37. Witt M, Carlson J, Brandt S, Cragin M. Constructing Data Curation
Profiles. International Journal of Digital Curation. 2009;4: 93--103.

% Please keep the abstract below 300 words

% Please keep the Author Summary between 150 and 200 words
% Use first person. PLOS ONE authors please skip this step.
% Author Summary not valid for PLOS ONE submissions.

% For figure citations, please use "Fig." instead of "Figure".

%\section*{References}
% Either type in your references using
% \begin{thebibliography}{}
% \bibitem{}
% Text
% \end{thebibliography}
%
% OR
%
% Compile your BiBTeX database using our plos2015.bst
% style file and paste the contents of your .bbl file
% here.
%

\end{document}